Millions of Type Errors

### 2023-08-11 programming23/paper.tex
quick analysis notes from gdoc

>>>>

\subsection{Analysis 2}

TODO work this into the above or below.

What can we learn from the telemetry? We have 1.5 million records from Feb 22
to April 14 (roughly 27K records per day).
They mention +70 million type errors.
(Some records point to a large number of errors. FILL median, max. Could be cascading,
which makes the 70million number uninterpretable.)
What takeaways are there??!

First,
users rarely switch type analysis mode. There are 3 modes available and users
are free to switch between them, or to mix modes in a codebase, but we rarely
detect these events:

\pct{99} of sessions use a single mode

\pct{0.2} of sessions use different modes in different modules. (More
precisely, the user switched to a module --- triggering a telemetry record ---
that used a different mode than the previous one.)

\pct{0.2} of records show a mode switch (equally split between upgrades and
downgrades, 341 and 320)

Second,
the stronger analysis modes are roughly 10x less popular than weaker ones
(across the \pct{99} of sessions that stick with one mode):

\pct{90} of sessions use nocheck mode
\pct{9.8} use nonstrict mode
\pct{0.2} use strict mode

Third,
In the edit range, we know exactly which errors showed up. This leads to a few
takeaways, such as:

Only 3 sessions (out of +300K) hit the typechecker's internal limits on problem
size --- CodeTooComplex errors. Good!

About \pct{90} of type errors are probably due to typos (SyntaxError, UnknownSymbol,
UnknownProperty). These are not propoer type errors.

10 errors never showed up. These include DynamicPropertyLookupOnClassesUnsafe and
SwappedGenericTypeParameter

Fourth,
Codebase size and session length tend to follow a power law. 
One aspect of size for nonstrict sessions: the number of
lines in the edit range. Over 80 sessions have 20-line ranges, about 40 have a
30-line range, etc.

TODO must explain edit range carefully.

TODO must show full plot, zoomed plot side by side

Fifth,
Increases and decreases to the number of type errors are evenly balanced across
the data.
In strict mode, \pct{24} of records increase the number of type errors. \pct{26}
decrease. \pct{50} keep it constant.
In nonstrict mode, \pct{14} increase errors, \pct{15} decrease, and \pct{71}
keep it constant
This plot is for nonstrict. X-axis = time since start of session, cut off at 30
seconds. Y-axis = change in num errors since last record, cut off at +/- 100.

\begin{figure}[t]
  \includegraphics[width=0.8\columnwidth]{img/example-error-delta.png}
  \caption{Errors delta, nonstrict}
  \label{f:error-delta}
\end{figure}

Sixth, a mystery. May be a bug in Luau.
May be DM insensitivity in strict mode
(if DM has top type that would lead to errors.)
Context: each of the 3 modes runs 2 kinds of type analysis. One is the regular
type analysis. The other is Forced Strict (FS), which converts all dependencies
to strict mode. The reason for FS is that it sometimes gives better
autocomplete suggestions etc.
You'd expect that the number of type errors is always <= num FS errors. But we
found several cases where this is false!
The plot here "stacks" every strict session into bars. X-axis is a record index
within a session (bars are higher on the left because we have lots more short
sessions than long ones). Y-axis counts the number of records where num TE <= num FS
is true (blue) and false (yellow).

TODO clarify semantic x-axis order

%% FS has: 1 all dependencies in strict, 2 data-model awareness (DM turns into
%% type graph, ... initial DM, all properties go into a table type; DM -> folder
%% -> model -> part .. get dot-driven autocomplete) in strict, DM might have type
%% Any might have type Instance, the latter would mean more errors than FS

\begin{figure}[t]
  \includegraphics[width=0.8\columnwidth]{img/example-tefs.png}
  \caption{Are there fewer type errors than forcedstrict errors? Expect true always, but that is not the case.}
  \label{f:tefs}
\end{figure}
